---
title: "Gov 50 Final Project"
author: "Gowri Rangu"
description: "Did private defense funding and party identification influence the House of Representatives' 2013 vote to continue the National Security Agency's (NSA's) mass surveillance of phone behaviors?"
output:
  distill::distill_article:
    self_contained: false
---
## **Introduction**
*Did private defense funding and party identity influence the House of Representatives' 2013 vote to continue the National Security Agency's (NSA's) mass surveillance of phone behaviors?* This study will examine the extent, or lack thereof, to which the amount of funding received from the private defense industry and party affiliation impacted a Representative’s stance in voting for or against the NSA’s mass cellphone surveillance. I hypothesize that the amount of funding received had a larger effect than that of political affiliation on the voter’s decision regarding privacy. Because the vote took place during the Obama administration, party loyalties could have skewed the vote as representatives are more likely to support the agenda of their party, especially if there is a corresponding president in power. Furthermore, since representatives rely on political sponsorships to propel their campaigns or policies, they are more inclined to vote for the outcome that will preserve their funding. 

The sample in this study is the House of Representatives from the 113th United States Congress, and the unit of analysis is each representative that voted on the NSA amendment. There are two explanatory variables of interest that will be tested: the representative’s party affiliation and funding from the private defense industry. The outcome variable is whether or not the representative voted for or against the continuation of NSA surveillance on American citizens. These variables were taken from the actual vote on the amendment and compiled into a dataset by the nonprofit research organization MapLight. To provide support for my hypothesis, a stronger, positive relationship between campaign financing and votes for mass surveillance must be observed compared to the relationship between party identity and votes. On the other hand, a weak or neutral relationship between sponsorships and votes supporting the amendment and/or a stronger relationship between party identity and votes would provide evidence against my hypothesis. Thus, with a regression analysis of the count of votes for passing the NSA amendment on the private defense funding and party identity indicator variable, a positive, higher coefficient for the funding variable would support my hypothesis. 

```{r echo = FALSE}
library(tidyverse)
library(infer)
library(kableExtra)


votes <- read_csv("vote_nsa.csv") |>
  mutate(position = ifelse(phone_spy_vote == "continue_allowing_mass_surveillance", 1, 0),
         party_bi = ifelse(party == "R", 1, 0),
         money_thous = money/1000) |>
  drop_na(phone_spy_vote)
```
## **Data and Research Design**
The data analyzed in this study is derived from the original votes of the U.S. House of Representatives on the 2013 NSA Phone Spying amendment. Since the independent and dependent variables were measured at the same time - after the vote was finalized - this observational study is of a cross-sectional design. The two explanatory variables are the amount of money (in thousands of dollars) that each representative received from the defense industry, a discrete variable, and his/her respective party, a binary variable. Coding an individual as a `0` signified that he/she was a Democrat while a `1` indicated Republican.

```{r echo = FALSE}
votes |>
  ggplot(aes(x = money_thous, fill = as.factor(position))) + 
  geom_histogram(aes(y = ..count../sum(..count..)), position = "stack", col = "white") + 
  labs(x = "Funding Recieved (in thousands)",
       y = "Proportion of Voting Stances",
       fill = "Vote")  +
  scale_fill_manual(values = c("red3", "forestgreen"))
```
The graph above shows the distribution of the outcome variable, the votes, as a proportion against an increasing x-axis of campaign funding from intelligence and defense firms. The data appears to be more clustered on the lower end of the amount of funding received, in which there is a higher portion of votes against the NSA amendment that notably decreases as the x-axis increases. 

```{r echo = FALSE}
votes |>
  filter(position == 1)|>
  ggplot(aes(x = position, fill = as.factor(party))) + 
  geom_histogram(aes(y = ..count../sum(..count..)), party = "stack", col = "white") +
  labs(
    title = "Distribution of Votes in Support of NSA Amendment by Party",
    x = "", 
    y = "Proportion of Votes for Mass Surveillance", 
    fill = "Party"
  ) + scale_x_discrete(labels=c("continue_allowing_mass_surveillance" = "")) +
  scale_fill_manual(values = c("steelblue1", "indianred1"))
```

The second graph illustrates the proportion of voters in support of the amendment by party identity. Based on the histogram, the Republicans accounted for more than half of the votes in favor of passing the bill compared to their Democrat counterparts. 

## **Results**
Before conducting the regression analysis, the bivariate relationship between the continuous variable of the value of sponsorships received and the binary variable of voting stance is illustrated with the below boxplot. The graph plots the funding on the x-axis and the voting outcomes on the y-axis. 

```{r echo = FALSE}
votes |>
  ggplot(aes(x = money_thous, y = phone_spy_vote, color = factor(position))) +
  geom_boxplot() +
  guides(color = F) +
  scale_color_manual(values = c("red3", "forestgreen")) +
  scale_y_discrete(labels=c("continue_allowing_mass_surveillance" = "For", "rein_in_phone_dragnet" = "Against")) +
  labs(x = "Funding Received (in thousands)",
       y = "",
       title = "Votes on NSA Surveillance")
```

The boxplot shows that there is a greater range of values in funding for those who voted for NSA surveillance. Moreover, there are higher minimum, median, and maximum values of campaign financing for votes supporting the bill. 

### **Regression Analysis**
Results from the multivariate regression analysis are below. The multivariate analysis regresses the NSA amendment votes on the two explanatory variables - campaign financing and political affiliation. 

```{r echo = FALSE}


fit <- lm(position ~ party + money_thous, data = votes) 

tab <- summary(fit)$coefficients

row.names(tab) <- c("(Intercept)", "Party", "Defense Funding (in thousands)")
colnames(tab) <- c("Estimate", "Std. Error", "T-value", "P-value")

kbl(tab, digits = 3) |>
  kable_styling(bootstrap_options = "striped", font_size = 12, full_width = F, position = "left")

```

These regression results show that there is a positive, statistically significant correlation between a representative’s party and defense funding and his/her vote on phone security. 

Holding campaign funding at a constant value, the multivariate model suggests that, on average, a one-unit increase of a representative identifying as Republican is associated with a predicted increase of 0.130 in votes supporting the NSA’s continued mass surveillance. On the other hand, holding party identity at a constant value, a one-thousand dollar increase in the amount of funding received is associated with a predicted increase of 0.003 in votes for passing the amendment. Comparing these two findings, there is a larger correlation between party affiliation and voting is clearly stronger than funds and voting.  

### **Statistical Significance**
In the multivariate analysis, the associations between party identity and voting and between funding and voting are statistically significant at the test level α = 0.05 as both p-values are approximately 0. Although there is statistical significance for both correlations, the stronger, more notable relationship is with the explanatory variable of party affiliation rather than defense funding as the former has a higher coefficient of 0.130. 

### **Causality and Confounders**
The findings from the regression analysis cannot be interpreted causally since the explanatory variables were not randomly assigned to each representative and his/her vote. In other words, because there was no control and treatment in this study, an increase in the number of Republicans or in the amount of funding cannot in turn mean a causal increase for increases in the number of votes supporting the NSA. 

There are a few possible confounding variables that were excluded from this study due a lack of data available. One possible confounder is the representative’s policy and political alignment with President Barack Obama who was president the same year of the NSA amendment vote. This factor influences the funding variable as representatives with more party support from the President will most likely not need as much campaign financing than those without the President’s backing. Moreover, this variable could have affected the results of the vote as representatives with a significant sense of party loyalty could have casted a vote that aligned with the Obama administration. 

Another possible confounder in this study is the general political ideologies of the representative’s home state as it could have impacted a representative’s party affiliation when they ran for office and had an effect on the representative’s vote if they chose to align their stance on mass surveillance with that of their state. 

## **Conclusion**
The regression results demonstrate that an increase in the number of Republican representatives is positively and significantly correlated to an increase in the number of votes in favor of preserving phone spying from the NSA. This correlation is also stronger than the one between the amount of funding and votes. In a multivariate analysis that controls for any possible confounding variables, there would be a greater estimate of this association. Both explanatory variables have statistical significance, providing partial support for my initial hypothesis. However, since the correlation coefficient of party identification is greater than that of the funding variable, my hypothesis that the funding variable would have a stronger effect on the outcome variable is disproven. 

A limitation of this study is that the representatives who abstained from voting were not taken into account in the analysis. This study was unable to find any plausible explanation as to why certain representatives chose not to vote. The second limitation is this study is constricted to a specific instance of Congress - its vote on a bill in 2013. Thus, the results cannot be generalized to indicate any patterns about the subsequent voting behaviors of Congress. 

### **Potential Improvements**
Since this study takes data from only a single vote of a Congressional bill, analyzing more votes on similar issues regarding phone spying and surveillance could provide deeper insight on whether or not there are patterns in the voting behavior of Representatives over a period of time. Lastly, another improvement may be studying the representatives that abstained from voting over this amendment and similar bills. 



